Tropical Mathematics and the Lambda-calculus I

### 2023-11-30 CSL24/main.tex
commit

\author{Davide {Barbarossa}}{Dipartimento di Informatica - Scienza e Ingegneria, Universit\`a di Bologna, %[optional: Address], 
Italy% \and My second affiliation, Country 
\and \url{https://lipn.univ-paris13.fr/\~barbarossa/index.html} }{davide.barbarossa@unibo.it}{https://orcid.org/0000-0003-4608-8282}{%(Optional)Funding each author
}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty. Please provide at least the name of the affiliation and the country. The full address is optional. Use additional curly braces to indicate the correct name splitting when the last name consists of multiple name parts.

\author{Paolo {Pistone}}{Dipartimento di Informatica - Scienza e Ingegneria, Universit\`a di Bologna, %[optional: Address], 
Italy% \and LIP, Universit\'e de Lyon, France 
\and \url{http://logica.uniroma3.it/pistone/
} }{paolo.pistone2@unibo.it}{https://orcid.org/0000-0003-4250-9051}{%(Optional)Funding each author
}

@@

\relatedversion{\url{https://arxiv.org/abs/2311.15704}} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversiondetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]{Classification (e.g. Full Version, Extended Version, Previous Version}{URL to related version} %linktext and cite are optional

@@

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Access}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\appendix

\section{Proofs from \autoref{section2}: Theorem \ref{theorem:fepsilon}}
\input{appSection3}

\section{Proofs from \autoref{section3}: $(\LREL, !_{n})$ is a model of $\BSTLC$}
\input{appSection5}

\section{Proofs from \autoref{sec:tls}: the Analysis of Tropical Power Series}\label{appsec:tls}
\input{app_section4}

\section{Proofs from \autoref{sec:TayLip}: Lipschitz meets Taylor}
\input{appSection7}

\section{Proofs of \autoref{sec:GMS}: $\Lawv$-modules and Generalized Metric Spaces}
\input{appSection8}

>>>>

\author{Davide {Barbarossa}}{Universit\`a di Bologna, %[optional: Address], 
Italy% \and My second affiliation, Country 
\and \url{https://lipn.univ-paris13.fr/\~barbarossa/index.html} }{davide.barbarossa@unibo.it}{https://orcid.org/0000-0003-4608-8282}{%(Optional)Funding each author
}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty. Please provide at least the name of the affiliation and the country. The full address is optional. Use additional curly braces to indicate the correct name splitting when the last name consists of multiple name parts.

\author{Paolo {Pistone}}{Universit\`a di Bologna, %[optional: Address], 
Italy% \and LIP, Universit\'e de Lyon, France 
\and \url{http://logica.uniroma3.it/pistone/
} }{paolo.pistone2@unibo.it}{https://orcid.org/0000-0003-4250-9051}{%(Optional)Funding each author
}

@@

\relatedversion{An extended version is available at \url{https://arxiv.org/abs/2311.15704}.} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversiondetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]{Classification (e.g. Full Version, Extended Version, Previous Version}{URL to related version} %linktext and cite are optional

@@

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{Aniello Murano and Alexandra Silva} 
\EventNoEds{2} 
\EventLongTitle{32nd EACSL Annual Conference on Computer Science Logic (CSL 
2024)} 
\EventShortTitle{CSL 2024} 
\EventAcronym{CSL} 
\EventYear{2024} 
\EventDate{February 19--23, 2024} 
\EventLocation{Naples, Italy} 
\EventLogo{} 
\SeriesVolume{288} 
\ArticleNo{35}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%\appendix
%
%\section{Proofs from \autoref{section2}: Theorem \ref{theorem:fepsilon}}
%\input{appSection3}
%
%\section{Proofs from \autoref{section3}: $(\LREL, !_{n})$ is a model of $\BSTLC$}
%\input{appSection5}
%
%
%\section{Proofs from \autoref{sec:tls}: the Analysis of Tropical Power Series}\label{appsec:tls}
%\input{app_section4}
%
%\section{Proofs from \autoref{sec:TayLip}: Lipschitz meets Taylor}
%\input{appSection7}
%
%
%\section{Proofs of \autoref{sec:GMS}: $\Lawv$-modules and Generalized Metric Spaces}
%\input{appSection8}

### 2023-11-30 CSL24/section2bis.tex
commit

%\begin{wrapfigure}{r}{0.5\textwidth}%
\begin{figure}
\begin{center}
\begin{tikzpicture}[scale=0.9]
\begin{axis}[samples=250]
\addplot[yellow,domain=0:0.8] {1+0.02};

@@

\begin{remark}\label{rmk:val_trop}
 There is a well-known relation between tropical polynomial/power series and usual polynomials/power series. 
%Given $Q,L$ semirings with units and zeros, $L$ commutative idempotent, if $L$ is totally ordered by the partial order: $\alpha \preceq \beta$ iff $\alpha +  \beta = \beta$, then one defines a \emph{valuation} \cite{Izhakian2015} to be a map $\mathrm{val}:Q\to L$ s.t.\ $\mathrm{val}(0)=0$, $\mathrm{val}(1)=1$, $\mathrm{val}(ab)=\mathrm{val}(a)\mathrm{val}(b)$, $\mathrm{val}(a+b)\preceq \max\set{\mathrm{val}(a),\mathrm{val}(b)}$.
%For example, the valuation which gives $1\in L$ on all the invertibles of $Q$ and $0$ otherwise, is called the \emph{trivial valuation} $\mathrm{val}_1$.
%Notice that if $\mathrm{val}:Q\to ([0,+\infty],+,\cdot,0,1)$ is a valuation, then $\mathrm{logval}(x):Q\to ([-\infty,+\infty],\min,+,+\infty,0)$, $\mathrm{logval}(x):=-\log \mathrm{val}(x)$ is a valuation.
For a power series $f(x)=\sum_{i\in I} a_i x^i$ (polynomials being the case for $I$ finite) defined on $[0,1]$ and coefficients $a_i$ in $[0,1]$, %and for a fixed valuation $\mathrm{val}:Q\to \Lawv$, 
one defines the \emph{tropicalisation} $\trop f:\Lawv \to \Lawv$ of $f$ as the tropical polynomial/power series function $\trop f(\alpha):=\inf_{i\in I} \set{-\log(a_i)+i\alpha}$.
%Remark that tropicalising w.r.t.\ the trivial valuation gives $\trop^{\mathrm{val}_1} \sum_{i\in I} a_i x^i = \inf_{i\in I} \set{i\alpha} = \trop^{\mathrm{val}} \sum_{i\in I\subseteq\N} x^i$ for all valuation $\mathrm{val}$, because in $\Lawv$ the unit is $0$. %and the zero is $\infty$.
%Remember that a_i \neq 0 by def of formal power series
%That is, it corresponds to quotienting the coefficients in $Q$ for the idempotency of $+$.
\end{remark}

@@

For instance, by looking at Fig~\ref{fig:plot1} it appears that, \emph{far from $0$}, $\varphi$ behaves like some of the polynomials $\varphi_{n}$.
In particular, %for all $\epsilon >0$, $f$ coincides on $[\epsilon,\infty]$ with some polynomial $\varphi_{n}$. More, precisely, 
$\varphi$ coincides on $[\epsilon,\infty]$ with $\varphi_{n}$,
for $\epsilon \geq 2^{-(n+1)}$ (the smallest tropical root of $\varphi_{n}$).
However, at
%
%$x\in [\epsilon,\infty]$, with $\epsilon>0$
%It can be proven by hand that $\varphi(x)$ is a $\min$ for all $x>0$.
 $x=0$ we have that $\varphi(x=0)=\inf_{n\in\N} 2^{-n}=0$, and this is the only point where the $\inf$ is \emph{not} a $\min$.
Also, while the derivative of $f$ is bounded on all $(0,\infty)$, for $x\to 0^+$ it tends to $\infty$.
In fact, this is a general phenomenon, as showed below:
%shared by all tps with a finite number of variables, i.e.~of the form $f(x)=\inf_{n\in \mathbb N^{k}}\{nx + \matr f(n) \}$, where 
%$x\in \Lawv^{k}$, $n\in \mathbb N^{k}$, $nx$ is the scalar product and $\matr f\in \Lawv^{{\mathbb N}^{k}}$ is a vector of coefficients. 
% 
%%This phenomenon is reminiscent of \cite[Example 7]{Ehrhard2005},
%%%Differentials and Distances in Probabilistic Coherence Spaces. FSCD 2019
%%which actually motivated our first investigations.
%In fact, this behavior is shared by all tps with \emph{finitely} many variables, as shown by the following result (we identify $!\set{1,\dots,k}\simeq \N^k$, so the matrix of a tps $f$ with finitely many variables $x=x_1,\dots,x_k$ (and one output variable) is given as a $\matr f:\N^k\to\Lawv$, $f$ having shape $f(x)=\inf_{n\in \N^k}\set{nx+\matr f(n)}$, with $nx$ the scalar product).

>>>>

%\begin{wrapfigure}{r}{0.5\textwidth}%
\begin{figure}
\begin{center}
\begin{tikzpicture}[scale=0.7]
\begin{axis}[samples=250]
\addplot[yellow,domain=0:0.8] {1+0.02};

@@

\begin{remark}\label{rmk:val_trop}
 There is a well-known relation between tropical polynomial/power series and usual polynomials/power series. 
%Given $Q,L$ semirings with units and zeros, $L$ commutative idempotent, if $L$ is totally ordered by the partial order: $\alpha \preceq \beta$ iff $\alpha +  \beta = \beta$, then one defines a \emph{valuation} \cite{Izhakian2015} to be a map $\mathrm{val}:Q\to L$ s.t.\ $\mathrm{val}(0)=0$, $\mathrm{val}(1)=1$, $\mathrm{val}(ab)=\mathrm{val}(a)\mathrm{val}(b)$, $\mathrm{val}(a+b)\preceq \max\set{\mathrm{val}(a),\mathrm{val}(b)}$.
%For example, the valuation which gives $1\in L$ on all the invertibles of $Q$ and $0$ otherwise, is called the \emph{trivial valuation} $\mathrm{val}_1$.
%Notice that if $\mathrm{val}:Q\to ([0,+\infty],+,\cdot,0,1)$ is a valuation, then $\mathrm{logval}(x):Q\to ([-\infty,+\infty],\min,+,+\infty,0)$, $\mathrm{logval}(x):=-\log \mathrm{val}(x)$ is a valuation.
For a power series $f(x)=\sum_{i\in I} a_i x^i$ (polynomials being the case for $I$ finite) defined on $[0,1]$ and coefficients $a_i$ in $[0,1]$, %and for a fixed valuation $\mathrm{val}:Q\to \Lawv$, 
one can define the \emph{tropicalisation} $\trop f:\Lawv \to \Lawv$ of $f$ as the tropical polynomial/power series function $\trop f(\alpha):=\inf_{i\in I} \set{-\log(a_i)+i\alpha}$.
%Remark that tropicalising w.r.t.\ the trivial valuation gives $\trop^{\mathrm{val}_1} \sum_{i\in I} a_i x^i = \inf_{i\in I} \set{i\alpha} = \trop^{\mathrm{val}} \sum_{i\in I\subseteq\N} x^i$ for all valuation $\mathrm{val}$, because in $\Lawv$ the unit is $0$. %and the zero is $\infty$.
%Remember that a_i \neq 0 by def of formal power series
%That is, it corresponds to quotienting the coefficients in $Q$ for the idempotency of $+$.
\end{remark}

@@

For instance, by looking at Fig.~\ref{fig:plot1} it appears that, \emph{far from $0$}, $\varphi$ behaves like some of the polynomials $\varphi_{n}$.
In particular, %for all $\epsilon >0$, $f$ coincides on $[\epsilon,\infty]$ with some polynomial $\varphi_{n}$. More, precisely, 
$\varphi$ coincides on $[\epsilon,\infty]$ with $\varphi_{n}$,
for $\epsilon \geq 2^{-(n+1)}$ (the smallest tropical root of $\varphi_{n}$).
However, at
%
%$x\in [\epsilon,\infty]$, with $\epsilon>0$
%It can be proven by hand that $\varphi(x)$ is a $\min$ for all $x>0$.
 $x=0$ we have that $\varphi(x=0)=\inf_{n\in\N} 2^{-n}=0$, and this is the only point where the $\inf$ is \emph{not} a $\min$.
Also, while the derivative of $f$ is bounded on all $(0,\infty)$, for $x\to 0^+$ it tends to $\infty$.
In fact, this is a general phenomenon, as showed below:
%shared by all tps with a finite number of variables, i.e.~of the form $f(x)=\inf_{n\in \mathbb N^{k}}\{nx + \matr f(n) \}$, where 
%$x\in \Lawv^{k}$, $n\in \mathbb N^{k}$, $nx$ is the scalar product and $\matr f\in \Lawv^{{\mathbb N}^{k}}$ is a vector of coefficients. 
% 
%%This phenomenon is reminiscent of \cite[Example 7]{Ehrhard2005},
%%%Differentials and Distances in Probabilistic Coherence Spaces. FSCD 2019
%%which actually motivated our first investigations.
%In fact, this behavior is shared by all tps with \emph{finitely} many variables, as shown by the following result (we identify $!\set{1,\dots,k}\simeq \N^k$, so the matrix of a tps $f$ with finitely many variables $x=x_1,\dots,x_k$ (and one output variable) is given as a $\matr f:\N^k\to\Lawv$, $f$ having shape $f(x)=\inf_{n\in \N^k}\set{nx+\matr f(n)}$, with $nx$ the scalar product).

### 2023-11-30 CSL24/section2bisbis.tex
commit

This way, the probabilistic evaluation of $M$ is presented as a \emph{hidden Markov model} \cite{Baum1966}, a fundamental statistical model, and notably one to which tropical methods are generally applied \cite{Pachter2004}.
A natural question in this case would be, for a fixed $\omega_0$:
%
%The tropical point of view allows now to express two natural questions about this situation:
%\begin{enumerate}
% \item What is the \emph{maximum likelihood estimator} for the event ``$M\twoheadrightarrow_{\omega_0} \false$''  (similarly for $\true$)?
% I.e., which is the choice of $p,q$ that maximizes the probability $P_{\omega_0}$?
% \item 
knowing that $M$ reached a normal form, say $\true$, %produced $\false$ (similarly for $\true$),
 what is the \emph{maximum likelihood estimator} for the event ``$M\twoheadrightarrow_{\omega_0}\true$''? %, knowing that ``$M\twoheadrightarrow \false$''?
In other words, what is the choice of $p,q$ that maximizes the conditional probability $\BB P(M\twoheadrightarrow_{\omega_0} \true\mid M\twoheadrightarrow \true)$, i.e.~that 
 makes $\omega_0$ the most likely path among those leading to $\true$?
% \mid ``M\twoheadrightarrow \false'')$)?
%\end{enumerate}
%
%Answering 1) amounts at solving a constrained maximization problem:
%\[(p,q)\in\arg\max_{y=1-x} P_{\omega_{0}}(x,y)\] or, equivalently: $p\in\arg\max_{x\in[0,1]} P_{\omega_0}(x,1-x)$,
%related to the monomial $P_{\omega_{0}}(p,q)$ and for which a tropical transformation does not, in general, help factoring the contraint $q=1-p$ out of the maximization.
%We will, instead, mainly concentrate on 2).

@@

@@

let us suppose that we observed the event ``$M\twoheadrightarrow\true$'', and thus our probabilities are conditioned under this observation. We have

@@

The tropical roots of $\trop Q_{\true}(x,y)$ are $(0,0)$ and all points of the form $(x,\frac{2}{3}x)$. Recall that these are the points where \autoref{eq:troproot} is satisfied for \emph{at least two} distinct values of $\omega_{0}$. From this it follows that $\trop Q_{\true} (x,y)=\trop P_{rrr}(x,y)=3y$ holds iff $y\leq \frac{2}{3}x$, and  $\trop Q_{\true}(x,y)=2x=\trop P_{ll}(x,y)$ otherwise.

@@

This means that, if $p\in[0,1]$ satisfies $1-p\geq p^{\frac{2}{3}}$ (for example, $p=\frac{1}{4}$), then $P_{rrr}(p,1-p) = \max_{\omega=ll, rll, rrr} P_{\omega}(p,1-p)$, i.e.: knowing that $M$ sampled $\true$ in its normal form, the most likely sampled occurrence of $\true$ is the one at the address $rrr$ iff $1-p\geq p^{\frac{2}{3}}$.

>>>>

This way, the probabilistic evaluation of $M$ is presented as a \emph{hidden Markov model} \cite{Baum1966}, a fundamental statistical model, and notably one to which tropical methods are generally applied \cite{Pachter2004}.
Then, a natural question in this case, for a fixed $\omega_0$, is the following:
%
%The tropical point of view allows now to express two natural questions about this situation:
%\begin{enumerate}
% \item What is the \emph{maximum likelihood estimator} for the event ``$M\twoheadrightarrow_{\omega_0} \false$''  (similarly for $\true$)?
% I.e., which is the choice of $p,q$ that maximizes the probability $P_{\omega_0}$?
% \item 
knowing that $M$ reached a normal form, say $\true$, %produced $\false$ (similarly for $\true$),
 what is the \emph{maximum likelihood estimator} for the event ``$M\twoheadrightarrow_{\omega_0}\true$''? %, knowing that ``$M\twoheadrightarrow \false$''?
In other words, what is the choice of $p,q$ that maximizes the conditional probability $\BB P(M\twoheadrightarrow_{\omega_0} \true\mid M\twoheadrightarrow \true)$, i.e.~that 
 makes $\omega_0$ the most likely path among those leading to $\true$?
% \mid ``M\twoheadrightarrow \false'')$)?
%\end{enumerate}
%
%Answering 1) amounts at solving a constrained maximization problem:
%\[(p,q)\in\arg\max_{y=1-x} P_{\omega_{0}}(x,y)\] or, equivalently: $p\in\arg\max_{x\in[0,1]} P_{\omega_0}(x,1-x)$,
%related to the monomial $P_{\omega_{0}}(p,q)$ and for which a tropical transformation does not, in general, help factoring the contraint $q=1-p$ out of the maximization.
%We will, instead, mainly concentrate on 2).

@@

\tag{$\star$}

@@

let us suppose that we observed the event ``$M\twoheadrightarrow\true$'', so that our probabilities are conditioned under this observation. We have

@@

The tropical roots of $\trop Q_{\true}(x,y)$ are all the points of the form $(x,\frac{2}{3}x)$. Recall that these are the points where \eqref{eq:troproot} is satisfied for \emph{at least two} distinct values of $\omega_{0}$ (indeed for $\omega_{0}\in \{ll, rrr\}$). From this it follows that $\trop Q_{\true} (x,y)=\trop P_{rrr}(x,y)=3y$ holds iff $y\leq \frac{2}{3}x$, and  $\trop Q_{\true}(x,y)=2x=\trop P_{ll}(x,y)$ otherwise.

@@

This means that, if $p\in[0,1]$ satisfies $1-p\geq p^{\frac{2}{3}}$ (for example, $p=\frac{1}{4}$), then $P_{rrr}(p,1-p) = \max_{\omega=ll, rll, rrr} P_{\omega}(p,1-p)$. In other words, knowing that $M$ sampled $\true$ in its normal form, the most likely sampled occurrence of $\true$ is the one at the address $rrr$ iff $1-p\geq p^{\frac{2}{3}}$.

### 2023-11-30 CSL24/sectionLaw.tex
commit

\subsection{$\Lawv$-Modules and Complete $\Lawv$-Categories}

@@

Moreover, a $\Lawv$-enriched \emph{functor} between $\Lawv$-categories is nothing but a non-expansive map $f:X\to Y$, since functoriality reads as $Y(f(x),f(y))\leq X(x,y)$.
Functors of shape $\Phi: X\times Y^{\op}\to \Lawv$ are called \emph{distributors} and usually noted $\Phi: Y \pfun X$. Notice that distributors $\Phi: Y\pfun X$ and $\Psi: Z\pfun Y$ can be composed just like 
ordinary matrices in $\LREL$: $\Psi\diamond \Phi : Z\pfun X$ is given by
$(\Psi\diamond \Phi)_{z,x}=\inf_{y\in Y}\Psi(z,y)+\Phi(y,x)$.

@@

\begin{definition}[weighted colimits]
Let $X,Y,Z$ be $\Lawv$-categories,
$\Phi: Z\pfun Y$ be a distributor and  $f:Y\to X$ be a functor.
A functor $g:Z\to X$ is the \emph{$\Phi$-weighted colimit of $f$ over $X$}, noted $\colim(\Phi,f)$, if for all $z\in Z$ and $x\in X$
\begin{align}
X(g(z),x)= \sup_{y\in Y}\left\{X(f(y),x)\menus \Phi(y,z)\right\}
\end{align}

@@

Observe that the usual Cauchy completeness for a $\Lawv$-category $X$ follows from its cocompleteness. Indeed, a Cauchy sequence $(x_{n})_{n\in \mathbb N}$ in $X$ yields two adjoint distributors $\phi^{*}:1\pfun X$ and $\phi_{*}:X\pfun 1$, where 
$\phi^{*}(x')=\lim_{n\to \infty}X(x_{n},x')$ and
$\phi_{*}(x')=\lim_{n\to \infty}X(x',x_{n})$. Hence,  
$\colim(\phi^{*},1_{X}):1\to X$ must be a point $x$ satisfying $0=X(x,x)=\sup_{y\in X}\lim_{n\to \infty}(X(y,x)\menus X(x_{n},y))$, which implies $\lim_{n\to \infty}X(x_{n},x)=0$.

@@

It turns out that the notions of $\Lawv$-module and cocomplete $\Lawv$-category are indeed equivalent. More precisely, the categories $\Mod$ and $\GMet$ are isomorphic \cite{Stubbe2006}. 
First, any $\Lawv$-module $(M,\preceq, \star)$ can be endowed with the structure of a $\Lawv$-category by letting
$M(x,y) = \inf\left\{ \epsilon \mid \epsilon \star x\geq y\right\}$. Moreover, a homomorphism of $\Lawv$-modules induces a cocontinuous functor of the associated $\Lawv$-categories. 
Conversely, in cocomplete $\Lawv$-categories it is possible to introduce a continuous $\Lawv$-action
via suitable weighted colimits called \emph{tensors} (cf.~\cite{Stubbe2014}):

\begin{definition}[tensors]
Let $X$ be a $\Lawv$-category, $x\in X$ and $\epsilon \in \Lawv$. The \emph{tensor of $x$ and $\epsilon$}, if it exists, is the colimit $\epsilon \otimes x:= \colim( [\epsilon],\Delta x)$, where
$[\epsilon]: \{\star\}\pfun \{\star\}$ is the constantly $\epsilon$ distributor
and $\Delta x:\{\star\}\to X$ is the constant functor. 
\end{definition}

>>>>

\subsection{$\Lawv$-Modules and Cocomplete $\Lawv$-Categories}

@@

Moreover, a $\Lawv$-enriched \emph{functor} between $\Lawv$-categories is nothing but a non-expansive map $f:X\to Y$, since functoriality reads as $Y(f(x),f(y))\leq X(x,y)$.
Functors of shape $\Phi:  Y^{\op}\times X\to \Lawv$ are called \emph{distributors} and noted $\Phi: X \pfun Y$. Two distributors $\Phi: Y\pfun X$ and $\Psi: Z\pfun Y$ can be composed just like 
ordinary matrices in $\LREL$: $\Psi\diamond \Phi : Z\pfun X$ is given by
$(\Psi\diamond \Phi)_{x,z}=\inf_{y\in Y}\{\Psi(y,z)+\Phi(x,y)\}$.

@@

\begin{definition}[weighted colimits]
Let $X,Y,Z$ be $\Lawv$-categories,
$\Phi: Z\pfun Y$ be a distributor and  $f:Y\to X$ be a functor.
A functor $g:Z\to X$ is the \emph{$\Phi$-weighted colimit of $f$ over $X$}, noted $\colim(\Phi,f)$, if for all $z\in Z$ and $x\in X$
\begin{align*}
X(g(z),x)= \sup_{y\in Y}\left\{X(f(y),x)\menus \Phi(y,z)\right\}
\end{align*}

@@

Observe that cocompleteness for a symmetric $\Lawv$-category $X$ implies the usual Cauchy completeness. Indeed, a Cauchy sequence $(x_{n})_{n\in \mathbb N}$ in $X$ yields two adjoint distributors $\phi^{*}:1\pfun X$ and $\phi_{*}:X\pfun 1$, where 
$\phi^{*}(x')=\lim_{n\to \infty}X(x',x_{n},)$ and
$\phi_{*}(x')=\lim_{n\to \infty}X(x_{n},x')$. Hence,  
$\colim(\phi^{*},1_{X}):1\to X$ must be a point $x$ satisfying $0=X(x,x)=\sup_{y\in X}\lim_{n\to \infty}(X(y,x)\menus X(y,x_{n}))$, which implies $\lim_{n\to \infty}X(x_{n},x)=0$.

@@

It turns out that the notions of $\Lawv$-module and cocomplete $\Lawv$-category are indeed equivalent. More precisely, the categories $\Mod$ and $\GMet$ are isomorphic \cite{Stubbe2006}. 
First, any $\Lawv$-module $(M,\preceq, \star)$ can be endowed with the structure of a $\Lawv$-category by letting
$M(x,y) = \inf\left\{ \epsilon \mid \epsilon \star x\geq y\right\}$. Moreover, a homomorphism of $\Lawv$-modules induces a cocontinuous functor of the associated $\Lawv$-categories. 
Conversely, in cocomplete $\Lawv$-categories it is possible to introduce a continuous $\Lawv$-action
via suitable weighted colimits called \emph{tensors}:
\begin{definition}[tensors, cf.~\cite{Stubbe2014}]
Let $X$ be a $\Lawv$-category, $x\in X$ and $\epsilon \in \Lawv$. The \emph{tensor of $x$ and $\epsilon$}, if it exists, is the colimit $\epsilon \otimes x:= \colim( [\epsilon],\Delta x)$, where
$[\epsilon]:1\pfun1$ is the constantly $\epsilon$ distributor
and $\Delta x:1\to X$ is the constant functor. 
\end{definition}
